---
title: "Make confusR data"
author: "David Lovell"
date: "16/06/2021"
output: html_document
bibliography: VRES.json
---

```{r setup, include=FALSE}
# Start with a clean environment
rm(list=ls())

knitr::opts_chunk$set(echo = TRUE)

library(tibble)
library(magrittr)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(jsonlite)
library(viridis)
```

#### About this document

This describes and generates the data used in the `confusR` package

# Example data
## binary.confusion
 
`binary.confusion` shows where the true/false positives/negatives of a confusion matrix go:

```{r make-binary.confusion, echo=FALSE}

matrix(
  c("TP", "FP",
    "FN", "TN"),
  nrow=2, 
  byrow = TRUE,
  dimnames = list(predicted=c("positive", "negative"), actual=c("positive", "negative"))
) -> binary.confusion
```

```{r comment=NA, echo=FALSE, eval=FALSE}
noquote(binary.confusion)
```

## Eddy and Eddy.equal
`Eddy` is a binary confusion matrix derived from Probabilistic reasoning in clinical medicine: Problems and opportunities [@tversky_probabilistic_1982]:

```{r make-Eddy, echo=FALSE}
Pos   <-    10     
Neg   <-   990
TPR   <-     0.792
FPR   <-     0.096
TP    <- round(Pos * TPR)
FN    <- Pos - TP
FP    <- round(Neg * FPR)
TN    <- Neg - FP

matrix(
  c(TP, FP,
    FN, TN),
  nrow=2, 
  byrow = TRUE,
  dimnames = list(predicted=c("cancer", "benign"), actual=c("cancer", "benign"))
) -> Eddy
```

```{r comment=NA}
Eddy
```

`Eddy.equal` is the confusion matrix we would expect if that classification system was applied to a sample of 1000 cases in which the prior probability of cancer was 50%:

```{r make-Eddy.equal, echo=FALSE}
Pos   <-   500     
Neg   <-   500
TPR   <-     0.792
FPR   <-     0.096
TP    <- round(Pos * TPR)
FN    <- Pos - TP
FP    <- round(Neg * FPR)
TN    <- Neg - FP

matrix(
  c(TP, FP,
    FN, TN),
  nrow=2, 
  byrow = TRUE,
  dimnames = list(predicted=c("cancer", "benign"), actual=c("cancer", "benign"))
) -> Eddy.equal

rm(Pos, Neg, TPR, FPR, TP, FN, FP, TN)
```


```{r comment=NA}
Eddy.equal
```

## Wiki

`Wiki` is a binary confusion matrix sourced from Wikipedia's [Likelihood_ratios_in_diagnostic_testing]( https://en.wikipedia.org/wiki/Likelihood_ratios_in_diagnostic_testing) in which there is a worked example of a diagnostic test with sensitivity 67% and specificity 91% applied to 2030 people to look for a disorder with a population prevalence of 1.48%

```{r make-Wiki, echo=FALSE}
matrix(
  c(20,  180,
    10, 1820),
  nrow=2, 
  byrow = TRUE,
  dimnames = list(predicted=c("cancer", "benign"), actual=c("cancer", "benign"))
) -> Wiki
```

```{r comment=NA}
Wiki
```

## CAMDA

`CAMDA` is a 23-class confusion matrix from a classifier we built for the 2020 CAMDA Challenge as described in our ISMB2020 conference paper [Metagenomic Geolocation using Read Signatures](https://www.iscb.org/cms_addon/conferences/ismb2020/tracks/camdacosi):


```{r read-CAMDA, echo=FALSE}
  read_csv("./data/CAMDA.csv", col_types = cols()) %$% 
  table(predicted, actual) %>%
  unclass() -> CAMDA
```

```{r comment=NA}
options(width = 120)
CAMDA
```

## CUP

`CUP` is the 17-class confusion matrix shown in Figure 2 of [AI-based pathology predicts origins for cancers of unknown primary](https://www.nature.com/articles/s41586-021-03512-4):

```{r read-CUP, echo=FALSE}


read_csv("./data/CUP.csv", col_types = cols())   %>%
  column_to_rownames('Organ') %>% 
  as.matrix() -> CUP
mode(CUP) <- "integer"
dimnames(CUP) <- list(predicted=substr(rownames(CUP), 1, 4), actual=substr(colnames(CUP), 1, 4))
```

```{r show-CUP, comment=NA}
options(width = 120)
CUP
```

## HASY

```{r}
HASY.names <- fromJSON("./data/hasy.json")
fromJSON("./data/confusion-matrix.json") %>% t() -> HASY
dimnames(HASY) <- list(predicted=HASY.names, actual=HASY.names)
```

```{r}
as.data.frame.table(HASY) %>% 
  as_tibble() %>%
  ggplot(aes(x=actual,y=predicted)) +
  geom_raster(aes(fill=Freq)) +
  #scale_fill_viridis() +
  scale_fill_gradient(low="white", high="black") +
  scale_y_discrete(limits=rev) +
  coord_equal() +  theme_void()
```


# Save data as RDS

```{r}
saveRDS(binary.confusion, "./rds/binary.confusion.rds")
saveRDS(Eddy,             "./rds/Eddy.rds")
saveRDS(Eddy.equal,       "./rds/Eddy.equal.rds")
saveRDS(Wiki,             "./rds/Wiki.rds")
saveRDS(CAMDA,            "./rds/CAMDA.rds")
saveRDS(CUP,              "./rds/CUP.rds")
saveRDS(HASY,             "./rds/HASY.rds")
```

Clear the environment and reload the data
```{r}
rm(list=ls())

readRDS("./rds/binary.confusion.rds") -> binary.confusion
readRDS("./rds/Eddy.rds")             -> Eddy
readRDS("./rds/Eddy.equal.rds")       -> Eddy.equal
readRDS("./rds/Wiki.rds")             -> Wiki
readRDS("./rds/CAMDA.rds")            -> CAMDA
readRDS("./rds/CUP.rds")              -> CUP
readRDS("./rds/HASY.rds")             -> HASY

ls.str()
```

# References
